### 2008-12-16 talk/ecoop2009/clibackend.tex
added picture with method and block ids

\commentout{

Using this approach, after a while the blocks of our original graph are scattered over a lot of different methods; however, there are no constraints about how these blocks can be linked together, so it happens to have links between blocks which are not in the same method. In the following, we will refer to them as external links.

If the external block we want to jump to happens to be at the beginning of its containing method, it is enough to invoke the method; but, what if we want to jump somewhere in the middle? What we really want is to produce a method which has multiple entry-points; again, doing it in assembly would be trivial, but the virtual machine does not provide any support for it, so we need a work around.

Each method in a graph is assigned an unique 16 bit method id; each block in a method is assigned a progressive 16 bit block number. From this two numbers, we can compute the block id as an unsigned integer, by storing the method id in the first 16 bits and the block number in the second 16 bits. By construction, the block id is guaranteed to be unique in the graph.

>>>>

What cannot be easily implemented in CLI is following an external link
whose target is not an initial block; consider, for instance, the
outgoing link of the block dynamically added in the right-hand side
picture of Figure~\ref{flexswitch-fig}. How it is possible to pass the
right arguments to the target block?

To solve this problem a special block, called \emph{dispatcher}, is
added to every method; whenever a method is invoked, its dispatcher is
executed first\footnote{Recall that the dispatcher is a special block
and must not be confused with the initial block of a method.} to
determine which block has to be executed.
This is done by passing to the method a 32 bits number, called 
\emph{block id}, which uniquely identifies the next block of the graph to be executed.
The high word of a block id is the id of the method to which the block
belongs, whereas the low word is a progressive number identifying
each block implemented by the method.

The picture in Figure~\ref{block-id-fig} shows a graph composed of three methods (for
simplicity, dispatchers are not shown); method ids are in red, whereas
block numbers are in black. 
The graph contains three external links; in particular, note the link
between blocks \texttt{0x00020001} and \texttt{0x00010001} which
connects two blocks implemented by different methods.
\begin{figure}[h]
\begin{center}
\includegraphics[height=6cm]{blockid}
\caption{Method and block ids.}\label{block-id-fig}
\end{center}
\end{figure}

\commentout{
Each method in a graph is assigned an unique 16 bit method id; each block in a method is assigned a progressive 16 bit block number. From this two numbers, we can compute the block id as an unsigned integer, by storing the method id in the first 16 bits and the block number in the second 16 bits. By construction, the block id is guaranteed to be unique in the graph.
